\chapter{POND \textit{in situ} Surveys}
\label{cha:app_POND_insitu_surveys}

\section{Weekly Survey}
Please answer these questions reflecting on your experience for the past week only. 

\noindent\textbf{Question 1.}
How challenging was tracking all of your food intake this week? 

\begin{itemize*}
\item	Very Easy
\item	Pretty Easy
\item	Fine; Neither easy or challenging
\item	Challenging
\item	Impossible
\end{itemize*}

\noindent\textbf{Question 2.}
How often did you look up food via search rather than use a +1 button? 

\begin{itemize*}
\item	Looked up a food via search all of the time
\item	Looked up foods most of the time
\item	Used +1 buttons most of the time
\item	Used a +1 button all of the time
\item	Other:  
\end{itemize*}

\noindent\textbf{Question 3.}
How much of the time did you make entries when appropriate? 

\begin{itemize*}
\item	I made entries every time I ate something that needed to be counted
\item	I made entries for most meals, about 2/3 of the time.
\item	I made entries about a third of the time, 1-2 meals per day
\item	I made occasional entries throughout the week
\item	I made no entries this week
\end{itemize*}

\noindent\textbf{Question 4.}
Which of the components below have been easy for you to track? 

\begin{itemize*}
\item	Dark Green \& Orange Vegetables
\item	Veggies
\item	Whole Grains
\item	Grains
\item	Fruit Juice
\item	Fruit
\item	Dairy
\item	Protein
\item	Oils
\item	Added Sugar
\item	Sodium
\item	Solid Fats
\end{itemize*}


\noindent\textbf{Question 5.}
What was easy about those components?  
 

\noindent\textbf{Question 6.}
Which of the components below have been challenging for you to track? 

\begin{itemize*}
\item	Dark Green \& Orange Vegetables
\item	Veggies
\item	Whole Grains
\item	Grains
\item	Fruit Juice
\item	Fruit
\item	Dairy
\item	Protein
\item	Oils
\item	Added Sugar
\item	Sodium
\item	Solid Fats
\end{itemize*}
\noindent\textbf{Question 7.}
What was challenging about those components?  
 

\noindent\textbf{Question 8.}
How certain are you that you will enter all appropriate foods even when: 

 	Very certain I will do it			Impossible for me to do	N/A
\begin{itemize*}
\item The food is not in the database	
\item You have to guess at the components \& values	
\item You have to read the \``More details'' box to figure out how to count it	
\item You forget to enter anything right after you eat	
\item You are with other people when you eat
\end{itemize*}

\noindent\textbf{Question 9.}
In the past week, how many times have you: 

\begin{itemize*}
\item not found a food in the database?		
\item had to guess at components \& values?		
\item had to read the "More details" to figure out a food's components?		
\item had to read the given "cheat sheet" (paper) to figure out a food's components?		
\item forgot to make an entry right after eating?		
\item been with other people when eating?		
\end{itemize*}

\noindent\textbf{Question 10.}
What are 3 challenges or frustrations you've encountered this week?  
 

\noindent\textbf{Question 11.}
What are 3 things you liked or been successful with this week?  
 

\section{Final Guided Interview Questions}
\begin{enumerate*}
\item How was your experience? [Give the person a chance to volunteer]
\item Were you able to make food entries right after you ate? When you were not able, why not?
\item Did you use the diary as often as you could? 
a.	What challenges you faced using the diary? 
\item Was there consistency in where or when you used the diary? 
\item What features did you wish for? 
\item Were there any food items you didn't want to record? Why?
\item Did keeping a food record change how you ate? How so?
\item How did you determine the amount of food you ate? How difficult was it for you to enter the amount of food that you ate at each meal?
\item Would you enjoy having a program such as this one on your cell phone? Why or why not?
\item Did you use the location tagging feature? 
\item Was there a pattern or a rule/guideline you used to decide when to search for a food rather than enter it via "+1"s? 
\item Did you ever refer to the extra information provided (the website/pdf)? 
How often? 
\item Was there a time when you didn't believe what the diary told you? 	Example?
\item Were you left feeling more or less capable of implementing these dietary change patterns in your life?  
\item What were 3 things you really liked about this diary? 
\item What were 3 things you really didn't like about this diary? 
\item Do you feel that our questions have covered your experiences with the diary and its approach? Is there anything else you find interesting, significant, or otherwise worthwhile to share? 

\item Potential other items: 
\begin{itemize*}
\item Tradeoff between "cost to enter" and "value of information"
\item Did you miss the detail that comes with entering food in a traditional food journal? 
\item Was the "time line" feature helpful/useful/used? 
\item Would you recommend it to friends? 
\item Who would you recommend this to?
\item Would you buy it? For .99? For 4.99? 
\item Did it make you more aware of processed versus non-processed foods? 
\item Did it make you more aware of what is in a food/how it's made? 
\item What were some of the rules you used to "enter a point"? (for example: one small apple is one point, a salad is 2 points, …)
\item How "normal" has the past 3 weeks been for you, apart from using this software? 
\item Did you use/change the goals? 
\item How has this experience compared to previous experience with food journaling? 
\item Sharing this data: would you be more or less willing to share this data with someone else: Doctor/health practitioner, coach/trainer/nutritionist, friend/family, social network. 
\item Goals: ``I achieved my goal of eating x number of <component> today!'', ``I achieved my <component> goal today!''
\item About how much of what you ate was journaled? 
\item How often did you remember to enter it when you ate it? 
\item Why did you choose to participate in this study?
\end{itemize*}
\end{enumerate*}

