\chapter{In-Lab Interface Comparison Surveys}
\label{cha:app_cont2_inlab_surveys}

\section{Initial Survey}
This was administered to all POND evaluation participants at the beginning of the in-lab study. 

\noindent\textbf{Question 1.}
Age: 


\noindent\textbf{Question 2.}
Gender: Male/Female

\noindent\textbf{Question 3.}
Occupation or Major: 

\noindent\textbf{Question 4.}
How often do you use a cell phone?
\begin{itemize*}
\item	Several times a day
\item	Daily (approx. 1-2 times)
\item	Weekly (approx. 1-2 times)
\item	Monthly (approx. 1-2 times)
\item	Almost never
\item	Never
\end{itemize*}

\noindent\textbf{Question 5.}
How often do you enter text on your phone (SMS, contact information, email, notes)?
\begin{itemize*}
\item	Several times a day
\item		Daily (approx. 1-2 times)
\item		Weekly (approx. 1-2 times)
\item		Monthly (approx. 1-2 times)
\item		Almost never
\item		Never
\end{itemize*}

\noindent\textbf{Question 6.}
What model cell phone do you have? 

\noindent\textbf{Question 7.}
Does it have a touch-screen?


\noindent\textbf{Question 8.}
How do you enter text on your cell phone?

\begin{itemize*}
\item	12-key keypad (traditional phone keypad)
\item	QWERTY keyboard
\item	Touch screen-- soft keyboard
\item	Other:  
\end{itemize*}

\noindent\textbf{Question 9.}
What are some of your health concerns or goals, in regards to nutrition?

\begin{itemize*}
\item	Lose weight
\item	Control portion sizes
\item	Eat "better"
\item	Eat less/more
\item	Other:  
\end{itemize*}

\noindent\textbf{Question 10.}
Have you ever used a food diary to track your food intake (eating) before?

\begin{itemize*}
\item	No
\item	Yes, on paper
\item	Yes, using software on a computer
\item	Yes, using a website on a computer
\item	Yes, using a PDA
\item	Yes, using a basic cell phone
\item	Yes, using a smart phone (iPhone, Windows Mobile, Symbian, etc)
\item	Other:  
\end{itemize*}

