\chapter{Background and Related Work}
\label{cha:relatedWork}

This chapter will cover related work, specifically work related to the self-monitoring of dietary intake. The process of self-monitoring dietary intake spans many fields: the medical community is concerned with any tools or processes that impacts a person's health; the nutrition and epidemiological community is concerned about how food impacts health, either individually or as a population; psychologists are concerned with the cognitive processes involved in both performing tasks and the higher level behavior change; and technology researchers are interested in how to design and build enabling technologies to support the process. The human-computer interaction researcher is concerned in spanning all of these fields to inform the design of the technology that ties it all together. 

For this dissertation, I focus on related work in nutrition and epidemiology, psychology and technology. My research goals relate to designing technology to support behavior change around nutrition behaviors, but research with the primary goal of collecting data via self-monitoring is helpful to relate to. Particularly in terms of beginning to identify what features of self-monitoring instruments have what goal, and characterizing how people use the different self-monitoring tools. 

One theme I see in the research is the distinction between self-monitoring of dietary intake to support behavior change versus as a data collection process. The distinction is that researchers know about the role of self-monitoring in the behavior change process in a number of domains. Research also makes a link between behaviors consistent with weight loss or weight maintenance. Weight loss and maintenance is concerned with monitoring calorie intake and expenditure, as well as a group of behaviors that positively correlate with good health outcomes. 


First, I'll talk about what self-monitoring of dietary intake is, and what is involved. Self-monitoring of dietary intake is the process that an individual uses to keep track of what they eat. The self-monitoring process is impacted by both the tool or instrument used for self-monitoring and the individual doing the self-monitoring. In regards to the individual, the process of self-monitoring involves cognitive processes, and the ability or willingness to continue self-monitoring involves individual goals, motivation, and resources.   The design of the tool impacts the process by clarifying (communicating? reminding?) goals as well as responding to individual motivation and resources. 

Then, I'll talk about why researchers are interested in both studying how people self-monitor dietary intake, as well as why some people may want to self-monitor dietary intake. 

We have three groups of researchers who are interested in studying the use of self-monitoring of dietary intake. These are epidemiologists and nutritionists; technology and human-computer interaction researchers; and psychologist and cognitive scientists. These three groups work together, but the research reflects boundaries between them. The research published in these different areas focus on different aspects of the self-monitoring problem. 

\section{Behavior Change}
In this section, I review some relevant theories in regards to behavior change. 
 
\subsection{Enabling change: Social Cognitive Theory}
Another relevant theory for PI designers and researchers is Social Cognitive Theory \citep{bandura_health_2004, bandura_self-efficacy:_1977}, which posits that a person뭩 behavior, environment and inner qualities all contribute to how a person functions. This theory has been applied to understanding how people learn, how social environments impact what people do, and how people regulate their own behavior. A key component in this theory is self-efficacy, which is summarized as a belief in one뭩 abilities.

Self-efficacy is traditionally measured by self-report. To develop self-efficacy measurements for a particular domain, researchers use open-ended approaches to identify common challenges and barriers to the problem. They then develop a series of statements of the form ``How confident are you that you can [achieve goal] even though [challenge]?'' with a 4-unit response scale ranging from ``Cannot do it'' to ``Highly certain can do''. An example of a statement is ``How confident are you that you can stick to a healthy eating plan after a long, tiring day at work?''

Research shows SE measures based on self-report indicate adherence to strategies to change behavior \citep{nothwehr_self-efficacy_2008}. While short-term studies cannot prove behavior change, SE measures provides valuable feedback about whether an intervention is supporting adherence to behavior change strategies, and indicate whether participants complete the study with an intention to continue.

This is an important feature for PI researchers: we are familiar with a domain and common challenges, so can build the scales easily; we usually use short-term studies to indicate long-term impact; and properly designed scales can help us to discover where a PI tool breaks down.

\subsection{Motivation for change: Transtheoretical Model}
The transtheoretical model (TTM) \citep{prochaska_transtheoretical_1997} reflects that people are in varying stages of change in relation to a given behavior. The stages include precontemplation, contemplation, preparation, action, maintenance and termination. 

\subsection{Supporting change: Goal setting and tending}
Here I talk about Locke and Latham \citep{locke_building_2002} and Consolvo and Landay \citep{consolvo_goal-setting_2009}. Goal theory. How to define goals. How to support progress toward goals. 

\subsection{Self-Monitoring and Change}
Research consistently supports that self-monitoring helps to mediate behavior change. Indeed, \citet{kanfer_self-monitoring:_1970} states ``self-observation is an initial step to self-directed behavior change''. Much research in this area falls outside the scope of this chapter, but in the domain of weight-loss, people who adhere to consistent self-monitoring of dietary intake lose more weight and have better indicators of related behaviors. Burke et al and the PREFER trial. WHI. \citep{greaves_systematic_2011}

It is unclear how self-monitoring impacts behavior change, or what mediates the process of self-monitoring. This is primarily due to limitations in the ability to evaluate the self-monitoring process. \citep{Baranowski1994, kanfer_self-monitoring:_1970, burke_self-monitoring_2005, glanz_improving_2006}. 

Later in this chapter I further explore how electronic self-monitoring tools for dietary intake are evaluated. 

\subsection{Self-Monitoring for Health}
Self-monitoring has been key to supporting lifestyle behavior changes necessary for treating cardiovascular disease, diabetes, cancer, and renal disease. 

Electronic self-monitoring has advantages over pencil and paper self-monitoring. 

Certain lifestyle behaviors can help to prevent disease as well as treat it. 

Since many of the behavior changes relate to nutrition and dietary intake, much research focuses on this. However, nutrition literature demonstrates that brief monitoring of the behaviors can still be effective. The technology domain has not put much emphasis on investigating brief monitoring instruments. 

Self-monitoring can help people make the behavior changes that prevent disease. 

The cognitive processes involved in the recall of food have been described in children as \citep{Baranowski1994}:
    Attention
    Perception
    Organisation
    Retention
    Retrieval
    Response formulation
People need to attend to the food that they are eating, at the time they eat it, in order to perceive it. The food information then needs to be organized and stored for retention. Then, the food information needs to be retrieved, and formulated into a response or action (e.g. creating a record). 

\subsection{Technology and Behavior Change (Persuasive Technology)}
In this section, talk about the body of work that has looked at how to use technology to support behavior change. Persuasive technology \citep{Fogg2002} and theory-driven technology \citep{consolvo_theory-driven_2009}. 

\section{Approaches to Changing Nutrition Behaviors}
Here I describe some standard approaches and theories nutritionists use to inform behaviors change to provide as background for reference later in this dissertation. It is not an exhaustive list. 

\subsection{Motivational Interviewing}
Motivational interviewing (MI) \citep{miller_motivational_2002, Miller2002} is a process practitioners use to identify what areas of change an individual is ready for. It is a technique employing goal setting, feedback mechanisms, and self-monitoring. There are three key elements: collaboration rather than confrontation; evoking ideas rather than imposing ideas; and autonomy rather than authority. In addition are four principles to follow: express empathy, support self-efficacy, roll with resistance, and ``develop discrepancy'' (or identify cognitive dissonance). 

MI has a strong record of impact for many types of behavior change \citep{Rubak2005}. It is particularly effective and well respected in the area of nutrition behavior changes \citep{abraham_taxonomy_2008, michie_effective_2009, greaves_systematic_2011}. 

\subsection{CBT Programs}
Cognitive-Behavioral Training (CBT) teaches people how to change the way they think about their nutrition behaviors to support them in the change. 

\subsection{Other techniques}
Other common themes in the research is regular group meetings, personalized feedback, nutritional education. 


\section{Related Work}
There are four projects closely related to the work presented in this dissertation. They are related because they are related to nutrition as opposed to physical activity, incorporate a form of self-monitoring, they take a user-centered design and evaluation approach, and they have some form of final validation. Work not covered here includes user-centered approaches to designing technology to support self-monitoring of physical activity, work focused primarily on developing or designing new technology to support self-monitoring of dietary intake, or work focused primarily on the evaluation of commercially available self-monitoring tools. However, I do refer to some of these projects later when I discuss the evaluation of tools that support self-monitoring of dietary intake. 

The four main projects related to this dissertation work include: Siek et al's \citep{siek_bridging_2009, siek_design_2006, siek_when_2006} work on self-monitoring of fluids and nutrients for patients with renal failure; Mamykina et al's work on supporting newly diagnosed diabetes patients; Tsai et al's work on mobile phone software that tracks both physical activity and dietary intake; and the Wellness Diary by Mattila et al, which is mobile phone software to track a variety of wellness factors, including dietary intake, physical activity, mood, stress, and sleep. 

Siek et al \citep{siek_bridging_2009, siek_design_2006, siek_when_2006} explored the use of PDA-based self-monitoring of dietary intake for a specific population: people who have chronic kidney disease (CKD) and require dialysis on a regular basis. People with CKD tend to have a number of conditions that make traditional food tracking more difficult. Dialysis patients must track their intake of certain nutrients (particularly sodium, potassium and fluids) in their  very carefully. Monitoring this information is particularly difficult for end-stage renal patients with poor literacy skills. The disease and treatment cause physical side effects that make using traditional electronic food journaling applications difficult. The physical side effects include degraded eyesight and fine muscle coordination, which makes small fonts or pointing targets inappropriate. To address these issues, the researchers investigated the use of barcode scanners and voice input to make food input easier. Over time participants preferred to use voice input because it was quicker and easier than using the scanner. 

Mamykina et al \citep{mamykina_examining_2011, mamykina_mahi:_2008} has explored supporting diabetes patients in managing their dietary intake. Specifically, MAHI focused on better supporting and exploring how to build tools to support people in their monitoring of a chronic disease, diabetes. Newly diagnosed diabetics usually go through a period of changing their dietary behaviors. MAHI is a mobile phone application that supports users in capturing and documenting eating episodes throughout the day. The records can then be reviewed later, encouraging reflection on the individual's behavior and choices. This can be characterized as a quick-capture with a strong emphasis on post-hoc analysis. Mamykina et al report \citep{mamykina_mahi:_2008} two specific items of interest for this dissertation: participants in particular had trouble using the Nokia N80 phones (including ``intimidat[ion] by the expensive-looking phone''), and there was a noticeable ``low level of personal interest'' in the use of the technology. Further work in this area focused more on social support and the building of community around diabetes management. 

Tsai et al developed the Patient-Centered Assessment and Counseling Mobile Energy Balance (PmEB) \citep{tsai_usability_2007},  an application that runs on a cell phone to help people self-monitor caloric balance in real-time.  The application on the phone allowed users to enter food intake and physical activity episodes, and displayed a visualization of current balance. They compared use of the application to self-monitoring with a paper diary. The comparison focused on characterizing the benefit of reminders in tracking caloric balance. Participants were assigned to one of three groups: paper (using a paper diary for tracking food intake and physical activity), one-reminder (using the phone application, with the phone providing one daily reminder to enter food), and three-reminder (again using the phone application, but with 3 daily reminders to enter food and activity). 

The Wellness Diary \citep{mattila_mobile_2008, ahtinen_user_2009, mattila_nuadu_2008, koskinen_customizable_2007, van_gils_feasibility_2001} is a  customizable diary software based on CBT. It provides many different wellness factors an individual can monitor over time. The wellness factors include quality of food intake, amount of physical activity, sleep, stress levels, and amount of time spent at work for the day. In regards to the self-monitoring of dietary intake, Mattila et al evaluated the use of the Wellness Diary by a self-selected group of people who were overweight and owned a phone that could run the software. This group was asked to use the Wellness Diary as part of a CBT-based weight management program. The group was provided with equipment and training necessary for them to implement the CBT-based weight management program. This included suggestions of what Wellness Diary measurements they should make (e.g. make food entries after meals, measure their weight every morning, etc). At the end of three months, participants were classified as in either a "LOSER" group (n=12), meaning they lost more than 1\% of their starting weight, or the "OTHER" group (n = 13). The participants in the LOSER group lost significantly more weight than the OTHER group (a loss of 2.94 kg versus a gain of 0.193 kg) and made significantly more entries per day (5.78 versus 5.24). Participants in the LOSER group made a mean of 3.41 food and drink entries per day, while those in the OTHER group only made 3.05. This study provides evidence to support previous work that indicate the longer a person continues their self-monitoring practices, the more weight they will lose (or the bigger the change). It also provides evidence that users appear willing to make five or six quick entries in a journaling application per day. 

The Siek and Mamykina work differ from the work presented in this dissertation in that they focus on a particular user population defined by disease. These user populations have well-defined constraints on dietary intake and are presumably more motivated to use technology to support their necessary self-monitoring. The Wellness Diary work differs in that it includes a wide range of behaviors to monitor. Self-monitoring of dietary intake is very cursory: each record captures whether the eating episode is health, not healthy, or unknown. PmEB is closely related to one of the projects I present in this dissertation, BALANCE. The primary difference is the evaluation: PmEB collected feedback on reminding strategies, while BALANCE looked at how well participants were able to make food entries. Both BALANCE and PmEB report challenges using a food database as part of the self-monitoring tool, which contributed to the pattern-based approach that informed the rest of this dissertation. 

Two other areas of related work that deserve a mention. One is that of using photography to capture a record of food eaten, detect what food was eaten, and calculate the nutrient values. The other is of using on-body sensors to detect eating episodes. One is attempting to automatically calculate food intake by observing the environment, while the other is attempting to automatically calculate food intake by observing the individual. 

The vision for using mobile-phone cameras for self-monitoring of dietary intake is a multi-step process. First, the user takes a picture of the food s/he intends to eat using the mobile phone. The process of taking the picture depends on which software is being used. The quality of the photograph, the lighting, the position of the camera all impact the ability of the software to automatically recognize the food. Some projects require the use of specific plates, while others use physical artifacts to determine camera angle and sizing information. Once the image is taken, computer vision strategies are used to identify the foods in the photo. Contextual information such as location, time and user may help the food identification process. When the user is done eating, s/he takes another photo of the food, showing the remaining food. This helps to identify how much food was eaten. A food database is used to look up the nutrition information and calculate calories consumed. 


FoodLog \citep{aizawa_food_2010, kitamura_foodlog:_2009, de_silva_clustering_2011} is a website that allows users to upload images of their meals. The images are analyzed using computer vision approaches to categorize food into five food groups (grains, vegetable, meat/beans, fruit, milk) \citep{kitamura_image_2010}. The system is primarily for personal use, but includes some features for users to connect and build community. It includes a number of analytical features and capabilities. Almaghrabi et al \citep{almaghrabi_novel_2012} uses before and after photos of food to identify the food, look up its nutritional content, and determine how much was eaten. It addresses the measurement problem by registering the image with the users thumb. It is restricted to meals eaten on white plates. DietCam \citep{fanyu_kong_dietcam:_2011, Kong2011} is a mobile-phone-based camera software that is exploring how to support users to take multiple photos of a meal, before and after eating, to improve the food recognition problem. Zhu et al \citep{zhu_use_2010, bosch_integrated_2011, zhu_image_2010, kim_development_2010} developed an image-based analysis system that consists of mobile-phone based image capture (of food), a database of known foods and nutrition information, and is working on the vision and classification aspects of this problem. DiaWare \citep{shroff_wearable_2008} supplements the image-capture and recognition problem with contextual information from the mobile phone, such as user, location, and time. 
PlateMate \citep{noronha_platemate:_2011} addresses the food recognition part of the problem by crowd-sourcing nutritional analysis from food photographs. Like The Eatery.  DietSense \citep{reddy_image_2007} is a project that automatically collects images all day, rather than initiated by a user. Initial studies explored if food images could be captured, identified, and analyzed to detect food intake. Finally, \citep{wen_wu_fast_2009} took videos of people eating in fast food restaurants, used image recognition to identify food items and how much was eaten, and used a nutrition database to calculate how many calories each person ate. 

All of these projects focus on the technology behind supporting the process of self-monitoring of dietary intake. The goal is to make the capture of eating episodes automatic, because so many people find it challenging to continue self-monitoring via self-report on a continual basis. Current challenges to using images for self-monitoring include food identification, determining the amount eaten, and collecting an appropriate photograph for processing. Collection is important because users have shown that they are as uncomfortable noticeably taking photos of their food for notation purposes as they are using a food diary in social situations. It is becoming more socially acceptable as more people are documenting food for social media. Automating the process will also help. 

Apart from the technical challenges, it is how the automatic capturing of nutritional behaviors impact the self-monitoring process. Self-monitoring consists of both a moment-of-capture and a moment-of-reflection, while automatic capture reduces this to the moment-of-reflection. Since it is unclear how the process of self-monitoring impacts behavior change, this is a potential area of concern and future research opportunity. 

Finally, image-based self-monitoring may have the same challenges as traditional, nutrient-focused self-monitoring. It is still unclear how well image-techniques identify nutrients such as added fat (e.g. butter), salt, or sugar. It also keeps the user's attention focused on a nutrient-oriented approach to dietary intake, as opposed to the pattern-based approach. I discuss this further in Chapter \ref{cha:app_cont2_inlab_surveys}. 

\citep{Pabler2011} and \citep{amft_detection_2005} are both working on sensing eating activities with the use of a wearable microphone. They are investigating whether features of the chewing sounds can determine which food is being chewed, and how much. 

WellNavi \citep{wang_development_2006} investigated the use of taking photos on a mobile phone to capture food intake. Participants took a picture of the food before and after eating.  A stylus-ruler was placed next to the meal to identify scale. Participants completed a short questionnaire about non-pictured foods. The images were then analyzed by an RD for nutritional content. There was little difference in the nutrition content as identified in the photos when compared to the weighed food records and 24-hr recall. This indicated that the use of photos with the extra data (ruler, survey) is comparable to existing standards of identifying nutritional intake. The record-capture process was reported as taking an average of 5 minutes. Participants reported being concerned about someone else seeing their photos. 


SapoFitness \citep{silva_sapofitness:_2011} presents the design and prototype of a physical activity and dietary intake diary. It is an Android application. In addition to tracking physical activity and dietary intake, features include a personal profile, feedback on current health status, and sharing. It also included predefined reminders to help the user make good choices throughout the day. 



Comber et al \citep{comber_supporting_2012} focused on using a mobile-phone based image capture system to track what patients in a hospital eat. This work is effective because it is used within a hospital environment where food served to patients is highly controlled, both in terms of content and serving size. 


Self-monitoring, defined as the process of observing and recording target behavior, has been identified as a key component of behavior change in general \citep{kanfer_self-monitoring:_1970} and for weight loss in particular \citep{michie_effective_2009} \citep{burke_effect_2011}. Of the people who have lost weight and kept it off (as registered with the Weight Loss Registry), the key behavior that correlates highly with losing and keeping weight off is the practice of self-monitoring both food (energy) intake and physical activity (energy expenditure) \citep{yon_personal_2007, burke_experiences_2009}. 

This led to the BALANCE project, where we combined physical activity detection with dietary intake monitoring to  encourage users to self-monitor. 

The research that indicates self-monitoring helps people to lose weight is not detailed enough to understand what aspects of self-monitoring are necessary. Additionally, this research is usually done on highly motivated participants. 

Please note: this is a very rich area for design and investigation. The research literature has many publications that are short, design-oriented, and not well evaluated. There are lots of great ideas out there. 


\section{Evaluation of Self-Monitoring}
So far I have presented background information to situate approaches to self-monitoring for behavior change, and described some related work that focused on the design and development of technology to support various aspects of self-monitoring. Now, I describe some of the evaluation approaches these projects used to properly situate the work I describe later in this dissertation. Understanding the how self-monitoring tools have been evaluated allows us to compare new work to previous work. 

\subsection{Challenges for Self-Monitoring}
The primary challenge of evaluating dietary intake self-monitoring tools  is that it is difficult for researchers to know how correct and complete the created records are. Methods researchers use to validated food diaries include food frequency questionnaires, 24-hour recalls, and doubly labeled water. Food frequency questionnaires ask people to report how often they eat specific foods in a given time period. Twenty four-hour recalls use a well-defined process elicit an accurate list of foods eaten in the past twenty-four hours. The doubly labeled water protocol uses isotopes to calculate an accurate base energy expenditure. The base energy expenditure, the reported calories consumed and body weight can be used to calculate the correctness of reported food consumption. Food frequency questionnaires are helpful for identifying general trends in a population, but a food record for a given time period may not be consistent with the food frequency questionnaire. Doubly labeled water is a resource intensive approach and has the drawback that it can only identify the number of calories consumed if body weight does not change. The calorie consumed calculation does not help identify the correctness of specific food diary entries.  Generally, the 24-hour recall is considered the best validation approach for evaluating the correctness of reported food intake. 

Overall, existing methods for evaluating self-monitoring and health outcomes considers the overall process of self-monitoring. They are not able to provide a more detailed analysis of what features of the self-monitoring tool are most or least effective for a given task or goal. 

\framebox{
\citep{junqing_shang_pervasive_2011} is developing the Dietary Data Recording System to both streamline the food data collection process and help evaluate self-monitoring technologies. It runs on an Android device and uses video and a laser grid to calculate food volume. 

\citep{wenyan_jia_food_2009} uses an LED (attached to a camera) to take a picture of food to calculate serving size. 

\citep{mela_honest_1997} describes how self-reports of food intake are ``Honest but Invalid''. }





\subsubsection{Nutrition and Epidemiological Research Domain}

The nutrition and epidemiological research areas include research to assess the process or outcomes of self-monitoring. There is less focus on the design or evaluation of the specific tool used for self-monitoring, and more focus on how that tool supports or does not support either the process or outcomes of self-monitoring. Research on the process of self-monitoring includes understanding how self-monitoring with various tools impacts the research process. More work focuses on outcomes of self-monitoring on larger goals, such as how self-monitoring impacts weight loss. 

The other concern of nutritionists is the task of collecting detailed information about a population's dietary intake to analyze it in terms of a number of variables (either descriptively, or to identify correlations between dietary factors and disease). Related research also includes work that is investigating how particular dietary changes impact disease; in this case, participants are required to self-monitor their dietary intake in order to confirm that they are following specific, prescribed nutritional goals. Researchers report on how different forms of self-monitoring support this larger research program. 

In the weight loss program research, research reporting on self-monitoring of dietary intake consists of studies where people participating in an existing weight loss and management program use electronic self-monitoring.  Examples of this are in the Women's Health Initiative \citep{glanz_improving_2006} and SMART trial \citep{burke_self-monitoring_2011}. Both programs incorporated electronic self-monitoring as part of a larger program that included a high level of support for participants. This support included technical support,  personalization of the software, and providing timely, personalized feedback on a regular basis.  Participants are already participating in a successful program consisting of education, group meetings on a regular basis, and receiving personalized feedback. In these studies, participants who do use electronic self-monitoring adhere to the program longer, and tend to lose more weight than participants who are not self-monitoring. Research in this area characteristically reports large number of participants monitoring for a long period of time (One month for \cite{glanz_improving_2006}, and six months for \cite{burke_self-monitoring_2011}). 

In the case of gathering detailed information about dietary intake for epidemiological purposes, the goal is to collect very correct data. The concern in this case is that the longer the time between eating and documenting, the greater the chance of error in the record. Therefore, the researchers want very timely records to improve the quality of the data. However, the populations are generally less motivated to document their food intake. It's also possible that the populations are less nutritionally literate, and usually don't have extensive training in the use of the monitoring instruments. Research that fits this profile tends to collect dietary intake information on the scale of a few days, rather than weeks. This puts the participants at a disadvantage, as dietary intake instruments usually require a few days (at least) to become familiar, and personalization or customization of the instrument doesn't result in much benefit. Research in this area characteristically reports many participants tracking for a short period of time (~3-7 days).

The final group of nutrition oriented research is concerned with the impact of various dietary recommendations over a long term. The Women's Health Initiative is an example. The Women's Health Initiative is tasked with studying the incidence of disease in women who have committed to following particular dietary guidelines, specifically low fat with high fruit, vegetable and whole grain intake. Women in this study need to document what they eat to provide verification to the program that they are following the diet. Here, the primary focus of the research is to support the participating women in following the diet, therefore the research program provides substantial resources and support to the participants, and additionally, the participants are highly motivated. Electronic self-monitoring has been studied in the context of making the recording process easier. Research in this area has the characteristics of large numbers of participants self-monitoring for a long period of time (1-6 months). 

In all of the above mentioned research, the self-monitoring instruments are frequently commercially available software, and when instruments are developed for the work, they are rarely described in detail. 

\subsubsection{Technology-oriented work}
As discussed earlier, technology research on self-monitoring focuses more on the design and development of technologies to support new self-monitoring technologies, or architecture to support self-monitoring. I am restricting this discussion to  work related to the mobile self-monitoring of dietary intake, although sometimes the work includes related behaviors such as physical activity or overall wellness. This discussion specifically excludes work that includes only a web-based component. 

\subsection{Evaluation Characteristics and Measures}
I identified and reviewed 15 studies reporting on an evaluation of self-monitoring for dietary intake \citep{glanz_improving_2006, burke_effect_2011,fukuo_development_2009,atienza_using_2008, yon_personal_2007, acharya_using_2011, arsand_usability_2007,jarvinen_hyperfit:_2008, mattila_mobile_2008, kozakai_dietary_2006, reddy_image_2007, silva_sapofitness:_2011, tsai_usability_2007, wang_development_2006, long_effectiveness_2012}. These studies reflect evaluations performed \textit{in situ}. I reviewed what instrument was used (commercially available or research prototype), the number of people, recruitment strategy or population, duration, the size of the database, and the goal of the evaluation. A summary of this information is presented in Table [ref].  

Published evaluations of mobile/electronic self-monitoring of dietary intake tools vary widely. HyperFit papers report 4 different evaluations, ranging from being used by individuals, to weight loss support groups, and by nutrition coaches, either by themselves or by clients. Most evaluations are 2-4 weeks, with a notable exception of WellnessDiary, which was used for 3 months by a population of people who wanted to lose weight. 

Apart from the work that focuses on a specific population, research in the technology domain mostly includes participants who are using the self-monitoring technology on their own. The population is frequently described as ``overweight individuals who want to lose weight'', and sometimes an education component is included, but for the most part, the study only looks at the use of the self-monitoring technology. This differs from the nutrition-oriented research that frequently has a substantial education and support component. 

One of the main differences we see between evaluations published in the nutrition versus technology domains is that 

One similarity we see in all of this work is that the process to capture detailed food and exercise records is frequently reported as ``too difficult''. This is partly due to general tediousness of creating entries, but partly due to the databases. 

Database size ranges from non-existent (Wellness Diary \citep{mattila_mobile_2008} uses a triage approach, specifying whether a meal is healthy, unhealthy or unknown, and \citep{atienza_using_2008} uses a 43-item survey to determine vegetable and whole grain intake) to very large (40,000-50,000) \citep{yon_personal_2007}. In all except one study it is noted that participants found it difficult to find or enter food into the system. SapoFitness \citep{silva_sapofitness:_2011} is the exception in that it does not provide enough detailed information about the evaluation.  Two systems reported pre-populating the mobile database with user-specified common foods \citep{tsai_usability_2007,glanz_improving_2006}, but participants still encountered foods not in the system. 

Evaluations vary widely as to whether participants carry their own device and install the software, versus being provided a device by the study. In the technology-oriented research, it is more common for participants to install the software on their own personal device, while in the nutrition studies devices are frequently provided. It may be more common for technology studies to provide the software for personal devices primarily because the research is more recent, and the availability of personally owned devices suitable for such software has increased substantially recently. This field may be more aware of good practices for ubiquitous computing evaluations. The nutrition literature, however, has been investigating this line of research for longer, and began when mobile devices were not as common in study populations. Additionally, for the research that is high impact, it may be important that everyone has the same software, which until recently has required very specific hardware. Also, for studies with many participants self-monitoring for a short time period, it is reasonable for the study to have a few devices for all participants to share. 

Reporting of measures varies depending on the goal and phase of the project. It would be helpful to be able to compare from study to study how the participants were using the software, to put it all in perspective. Usually an overall number is given: mean number of records per day. Sometimes rather than records, the number is the number of ``compliant'' days, or days when there are records entered at 3 different times. Usually there is a measure of adherence over time: the number of entered versus expected entries overall, as well as for the first week and the last week. 


\subsubsection{Usability and User Experience Evaluation of Food Diaries}


Food diaries are difficult to evaluate in lab settings due to the personal, situated nature of food consumption. As discussed above and in \citep{Baranowski1994}, three key steps in the process of self-monitoring of dietary intake are those of attention, perception and organization. The user must attend to the food intake, perceive the food, and observe the relevant features of the food. The steps of perceiving and organizing the food include recognizing what food it is, what extras may be included with it, and how much food is presented or consumed. In the real world, users are usually involved in the selection or preparation and consumption of the food, so have more context to be able to correctly identify and account for it. 

The goal of usability studies are to collect comparable data of users accomplishing well-defined tasks. This requires consistent, comparable, well-defined tasks for users to perform.  

Task presentation can be a textual list of well specified foods, with the food and serving sizes either exactly represented in the database or not. Exact specifications help to test a best-case scenario of finding and entering a food into the diary, while non-exact specifications reflect a more natural approach. However, this approach does not address the process of a user identifying the term to use to search the food database. An alternate approach to task presentation is to use photographs of food or real food. These presentation approaches are more like a real-world food specification task in that users need to generate the word to use to search the food database. However, food photographs have the downside of being challenging to identify serving sizes or added condiments. Overall, task presentation impacts which aspects of the food diary entry process a researcher can draw conclusions about. 

Task content is another area of concern in usability studies. Familiarity of foods vary from person to person. For some aspects of usability, food familiarity is not an issue.    A key challenge to finding a food in a food database is the nutritional literacy problem. 

One approach is to choose a few foods, specify exact amounts and look at how 

\textbf{In this section, talk about different ways to present foods input to a food diary, and the concerns about different approaches. To show real food? Do a 24-hr recall? etc. In Cont2, I'll refer back to this. }
One area of focus was the food tasks. Since we want to apply the findings of this study to the real world, one approach is to use a food diary or recall of that participant. This gives the benefit of a person being familiar with the food, and it would allow us as researchers to identify how the interfaces do or do not support the real world food. Familiarity is important, particularly with the index-based approach, because if people do not know what is in a food they are more likely to get it wrong. However, people have different foods they are familiar with, particularly in terms of content. Familiarity can depend on age, religion, geography, vegetarian or not, even gender. However, if each participant used different foods, we would not be able to compare timing and correctness measures for the different interfaces. Also, some participants might eat more ``single component'' foods while other participants eat more combination or restaurant foods, which are more challenging to score. Therefore, we decided it was best for all participants to use the same food tasks. 

Since familiarity is so important, we carefully considered how to choose the food for the tasks. One considered approach was to refer to a published gold standard. [blah blah, not published, not an issue in nutrition research, etc. ] We considered using published diet plans, such as those advocated by the American Heart Association for healthy diets. The drawback to those is that they are recommended, and may not accurately reflect the actual diet of the participants. Additionally, since the recommended menus are based on the official Food Pyramid, and one chosen food index is also based on the Food Pyramid, this could influence the correctness scores. After consideration, we decided to choose the food tasks based on the food diaries collected in the BALANCE focus groups. Since the study populations are drawn from the same underlying populations, the issue of familiarity is addressed. More details are explained later. 

Another important aspect is how the food tasks are presented to participants. In the real world, one of the important challenges to ``perfect'' food journaling is the process of seeing/eating food, identifying what it is, sometimes identifying methods of preparation (fried versus broiled), identifying un-seen characteristics (low-fat versus full-fat milk, with butter), and identifying portion or serving sizes in order to correctly enter the food (or choose from a list in a db). Sometimes this process is easy, as with packaged foods or when preparing your own food; other times such as at a restaurant it is more challenging. Literacy is also an issue: with a traditional journaling approach, it is one thing to be able to identify one cup of spaghetti, but if you are not able to spell it properly, you might not be able to find it in a database. Due to this process, we considered that using real plates of food would be more life-like, and that using photographs of food were close to real. However, with the use of photographs, care must be taken to obtain them properly. Scale and lighting are two important considerations. And, with research, we found that there exist food photograph booklets specially prepared for the purpose of studying people's estimation of portion sizes. However, these booklets had the familiarity concerns outlined above. We decided that portion size estimation and literacy were not primary concerns for this study, so chose to present food tasks as a list of foods on a card. 

\subsection{Evaluation of Electronic Self-Monitoring}
As addressed in \cite{klasnja_how_2011}, the evaluation of technology tools to support behavior change is difficult. This theme arose throughout the course of this dissertation work: what can we evaluate, how do we measure it, and what do we learn? In addition to evaluating this work individually, we wanted to be able to compare the results of our evaluation to previous work. This was more difficult than anticipated. Existing work evaluating self-monitoring of dietary intake is not consistent, and does not report results that can be used to compare one self-monitoring approach to another. 

Evaluation of technologies need to include both usability and utility: Usability reflects how usable the system is, while utility investigates the system's usefulness in a larger context.  

Given this challenge, this section focuses on the evaluation of self-monitoring of dietary intake tools. 
Much of literature in the nutrition community focuses on whether a tool in general improves the self-monitoring process. In the technology literature, evaluations focus more on documenting the design process and whether the technology works in regards to the intended impact. 






\section{Summary}In this section, I reviewed relevant background information for this dissertation, including behavior change theories, traditional approaches specifically for supporting nutrition behavior change, and the important role of  self monitoring. I provided some overviews of how self-monitoring has been studied both from a technology development perspective, and how nutritionists evaluate self-monitoring of dietary intake from a utility perspective. I also provided a short overview of projects related to those discussed in this dissertation. 
